Will Kamala Harris will win the 2024 Presidential election

### 2024-11-04 paper/paper.qmd
more caption changes

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-modelNPW
#| fig-cap: "Poll predictions of Nevada, Pennsylvania, Wisconsin"

@@

ggplot(subset_down, aes(x = end_date_num, y = median_prob, color = state)) +
  geom_line(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 10), se = FALSE) + 
  labs(
    x = "Days Since First Poll",
    y = "Predicted Probability"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ state, ncol = 2) +
    theme(legend.position = "none")

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-supporttable
#| tbl-cap: "Averaged predicted support for Harris by state"

@@

Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%. While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant Republican victory in 2016 and a rather marginal win in 2020, which was the first Democratic victory for the state since 1992. Georgia is in an interesting situation as lot of the Democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.

@@

Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of Democratic votes come from large cities and the majority of Republican votes coming from rural areas. Nevada has been a Democratic state for the last 4 elections with at least a 2% difference in favor of the Democrats in the last two elections.

Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the Democratic candidates. Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump. Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the Democrats.

@@

North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was Democratic and voted for Obama. This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the Republican with 49.8% of the vote, with 46.2% going to the Democrats. While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8%.

@@

Harris is currently leading the polls in Michigan with 49% of the vote and Trump with 47%, this is a rather slim margin, but the first in the discussed states where Harris is leading in the polls. In @fig-supportAM we can see that there was a significant jump for Harris and a small dip in October followed by a gradual increase leading up to the elections. When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner. Although unlike the previously mentioned states 2016 was the first time the state voted for the Republican candidate since 1992. Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote Democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.

@@

Second, Similarly to Nevada and Michigan Pennsylvania Voted for Trump in 2016 with a 48.2% to a 47.5% victory, and for Biden in 2020 with a slightly more significant but not overwhelming victory at 50% to 48.8%. 2016 was the first time Pennsylvania voted for a Republican candidate since 1988.

Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more Democratic than Republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.

@@

Without considering the swing states and if we assume all the other states will vote as expected Kamala Harris has a slight lead of 226 electoral votes to Donald Trump's 219. Now, if we take the 43 combined votes of Arizona, Georgia, and North Carolina, and add it to Trumps tally. While allocate Nevada, Michigan, Pennsylvania, and Wisconsin's votes to Harris which sum up to 50 electoral votes. This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes. Below we have added a makeup of what we believe the election to look like in an image @270towin:

@@

```{r, fig.width=5, fig.height=3}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ppcheck
#| fig-cap: "Posterior predictive check of model"

>>>>

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-modelNPW
#| fig-cap: "Predicted Probability of Support for Harris Over Time in Nevada, Pennsylvania, and Wisconsin Since Harris' Candidacy Declaration"

@@

ggplot(subset_down, aes(x = end_date_num, y = median_prob, color = state)) +
  geom_line(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs", k = 10), se = FALSE) + 
  labs(
    x = "Days Since First Poll",
    y = "Predicted Probability"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(hjust = 1)) +
  facet_wrap(~ state, ncol = 2) +
    theme(legend.position = "none")

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-supporttable
#| tbl-cap: "Average Predicted Support for Harris by Swing State"

@@

Considering the last two elections Georgia is very similar to Arizona, as in 2016 Trump won the state with 50.8% to Clinton's 45.6%. While in the 2020 election Biden won with a slim victory of 49.5% to Trump's 49.3%, which again is very similar to Arizona's last two election results with significant republican victory in 2016 and a rather marginal win in 2020, which was the first democratic victory for the state since 1992. Georgia is in an interesting situation as lot of the democratic vote in the state comes from the black community, although Harris a has less support from male black population, as the last victory for Biden was as a small margin, even a small percent of the population switching their vote could heavily impact the vote in 2024.

@@

Nevada is a state who's majority is dessert and wilderness, with two large cities being Reno and Las Vegas, as the case is with many states a Majority of democratic votes come from large cities and the majority of republican votes coming from rural areas. Nevada has been a democratic state for the last 4 elections with at least a 2% difference in favor of the democrats in the last two elections.

Interestingly enough, if we look at @fig-supportGNN we can see Nevada, similar to several other states had a spike in support of Harris in early September peaking north of 48% in the polls and their recent history of voting in favor of the democratic candidates. Nevertheless, Harris is seen to have had a gradual decline in the state stabilizing at around 46%, which which seems low in comparison to other states that are looking more favorably at Harris than Trump. Still, when we look at @fig-supportstate Harris is still has a marginal to Trump, considering all of these factors it is likely Harris will in fact take Nevada for the democrats.

@@

North Carolina has been an exclusively Republican state since 1980, with a singular exception in 2008 when the state was democratic and voted for Obama. This means that North Carolina voted for Trump in th last two elections, in 2016 there was more significant win for the republican with 49.8% of the vote, with 46.2% going to the Democrats. While in 2020, the Republicans maintained their support from the state but the margin of victory was slimmer with them receiving 49.9% of the vote and Biden getting 48.8%.

@@

Harris is currently leading the polls in Michigan with 49% of the vote and Trump with 47%, this is a rather slim margin, but the first in the discussed states where Harris is leading in the polls. In @fig-supportAM we can see that there was a significant jump for Harris and a small dip in October followed by a gradual increase leading up to the elections. When we look at the voting history of the state in the last two elections, the have voted for the same candidates as Arizona and Georgia, where they voted for Biden in 2020 and for Trump in 2016, with the voting percentages at 50.6% to 47.8% in 2020 and 47.5% to 47.3 in 2016 with respect to the winner. Although unlike the previously mentioned states 2016 was the first time the state voted for the republican candidate since 1992. Considering Michigan is home to the largest black majority city in the United states, Detroit, and a large majority of union workers who often vote democratic as it is seen as the more union-friendly party between the two it is rather like that Harris will win the state.

@@

Second, Similarly to Nevada and Michigan Pennsylvania Voted for Trump in 2016 with a 48.2% to a 47.5% victory, and for Biden in 2020 with a slightly more significant but not overwhelming victory at 50% to 48.8%. 2016 was the first time Pennsylvania voted for a republican candidate since 1988.

Lastly, if we consider @fig-supportWP we see that harris has been polling at a consistent 48% since mid september which is somewhat similar to Georgia, while Pennsylvania seems to be leaning more democratic than republican than Georgia.@fig-supportGNN, they have both similarly increased their interest in Harris since the polling began until mid september which is when they both becan to plateau at their current polling numbers.

@@

Without considering the swing states and if we assume all the other states will vote as expected Kamala Harris has a slight lead of 226 electoral votes to Donald Trump's 219. Now, if we take the 43 combined votes of Arizona, Georgia, and North Carolina, and add it to Trumps tally. While allocate Nevada, Michigan, Pennsylvania, and Wisconsin's votes to Harris which sum up to 50 electoral votes. This totals to a Democratic victory for Kamala Harris at 276 votes to Donald Trump's 262, we have come to the conclusion that Kamala Harris is likely to win the upcomming United States election by 14 electoral votes.

@@

```{r, fig.width=5, fig.height=3}
#| echo: false
#| warning: false
#| message: false
#| label: fig-ppcheck
#| fig-cap: "Posterior Predictive Check of Model"
